---
title: "Analytical Solution of the Neoclassical Growth Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analytical Solution of the Neoclassical Growth Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to solve the **deterministic neoclassical growth model** analytically, and how the results compare to the numerical solution computed by the `rcpp_neogro()` function.

The model assumes a representative household that chooses capital \(k_{t+1}\) to maximize:

\[
\sum_{t=0}^{\infty} \beta^t u(c_t),
\]
subject to:
\[
c_t + k_{t+1} = f(k_t) + (1-\delta)k_t,
\]
where:

- \(f(k) = k^\alpha\) is the Cobb–Douglas production function,
- \(u(c) = \frac{c^{1-\sigma} - 1}{1-\sigma}\) is the CRRA utility function,
- \(\beta \in (0,1)\) is the discount factor,
- \(\delta\) is the depreciation rate of capital.

# Steady State

In steady state, \(k_{t+1} = k_t = k^*\) and \(c_t = c^*\). The Euler equation is:

\[
u'(c_t) = \beta u'(c_{t+1}) [f'(k_{t+1}) + 1 - \delta].
\]

Using the CRRA utility function \(u'(c) = c^{-\sigma}\), this simplifies to:

\[
1 = \beta [\alpha (k^*)^{\alpha-1} + 1 - \delta].
\]

Solving for \(k^*\) yields:

\[
k^* = \left( \frac{\alpha}{\frac{1}{\beta} - 1 + \delta} \right)^{\frac{1}{1-\alpha}}.
\]

Once \(k^*\) is known, consumption \(c^*\) is:

\[
c^* = k^{*\alpha} - \delta k^*.
\]

# Analytical Computation in R

```{r}
# Parameters
beta  <- 0.96
alpha <- 0.28
delta <- 0.05
sigma <- 2.0

# Steady state capital
k_star <- (alpha / (1/beta - 1 + delta))^(1/(1 - alpha))

# Steady state consumption
c_star <- k_star^alpha - delta * k_star

k_star
c_star
```

# Comparison with Numerical Solution

The `rcpp_neogro()` function numerically solves the Bellman equation using value function iteration. We can check if the policy function converges to \(k^*\).

```{r}
library(neogro)

# Numerical solution
k_steady <- k_steady_state(beta = beta, alpha = alpha, delta = delta)

# Numerical solution
res <- neogro(NK = 400, beta = beta, alpha = alpha, delta = delta)

# Approximate steady state from the policy function
k_num <- res$k[which.min(abs(res$kopt - res$k))]

cat("Analytical k* =", k_star, "\n")
cat("k_steady k* =", k_steady, "\n")
cat("Numerical k*  =", k_num,  "\n")
```

# Conclusion

The analytical solution provides a benchmark for verifying numerical methods.  
For more complex versions of the growth model (e.g., stochastic shocks), the numerical approach becomes essential.
