---
title: "Simulating the Neoclassical Growth Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating the Neoclassical Growth Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

# Introduction

This vignette works through a fuller example than the quick start in the
README: we solve the model under a standard real-business-cycle calibration,
inspect the resulting policy and value functions, sanity-check that the
solver respects the resource constraint, and simulate a capital transition
path back to the steady state.

```{r setup}
library(neogro)
library(ggplot2)
```

# Solving the model

```{r}
alpha <- 0.36
delta <- 0.05
beta <- 0.96
sigma <- 1.0

k_steady <- k_steady_state(alpha = alpha, beta = beta, delta = delta)

policy <- neogro(
  NK = 501,
  alpha = alpha,
  beta = beta,
  delta = delta,
  sigma = sigma,
  TOL = 1e-5,
  TOL1 = 1e-7,
  verbose = FALSE
)

policy_file <- file.path(tempdir(), "policy.csv")
write.csv(policy, policy_file, row.names = FALSE)
policy <- read.csv(policy_file)

policy$resid_gdp <- policy$y - policy$copt - policy$iopt
```

# Policy, value, and residual diagnostics

```{r, fig.width=7, fig.height=5}
policy_long <- data.frame(
  k = rep(policy$k, times = 5),
  name = factor(
    rep(c("kopt", "copt", "iopt", "y", "resid_gdp"), each = nrow(policy)),
    levels = c("kopt", "copt", "iopt", "y", "resid_gdp")
  ),
  value = c(policy$kopt, policy$copt, policy$iopt, policy$y, policy$resid_gdp)
)

ggplot(policy_long, aes(k, value)) +
  geom_line() +
  geom_rug() +
  facet_wrap(~name, scales = "free_y") +
  labs(x = "Capital (k)", y = NULL)
```

`resid_gdp` should be numerically zero everywhere: it is the gap between
output and its uses (consumption plus investment), so this panel is a check
that the solver respects the economy's resource constraint.

# Investment vs. depreciation

```{r}
ggplot(policy, aes(k)) +
  geom_line(aes(y = iopt, color = "Investment")) +
  geom_line(aes(y = delta * k, color = "Depreciation (delta * k)")) +
  labs(x = "Capital (k)", y = NULL, color = NULL)
```

Where the two lines cross is (approximately) the steady state: investment
exactly offsets depreciation, so capital stops growing.

# Value function

```{r}
ggplot(policy, aes(k, vopt)) +
  geom_line() +
  labs(x = "Capital (k)", y = "Value function")
```

# Simulating a transition path

Starting below steady state, repeatedly applying the (interpolated) policy
function should converge capital toward `k_steady`.

```{r}
k0 <- 0.75 * k_steady
horizon <- 200

k_path <- numeric(horizon)
k_path[1] <- k0
for (t in 2:horizon) {
  k_path[t] <- approx(policy$k, policy$kopt, k_path[t - 1])$y
}

ggplot(data.frame(time = seq_len(horizon), k = k_path), aes(time, k)) +
  geom_line() +
  geom_hline(yintercept = k_steady, color = "red", linetype = "dashed") +
  labs(x = "Time", y = "Capital (k)")
```

# Conclusion

This worked example illustrates how to solve the model under a realistic
calibration, validate the solution against the economy's resource
constraint, and use the policy function to simulate capital dynamics away
from steady state.
